---
title: Quantifying the effects of individual player on goal scoring through a Bayesian approach
author: "Barba Paolo"
date: "July 3, 2023"
output:
  beamer_presentation:
    theme: "CambridgeUS"  # Choose a Beamer theme
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)
library(lme4)
library(R2jags)
library(ggplot2)
library(tidyverse)
load('Rdata/shots.RData')
```

## Introduction (1)

Nowdays, football clubs have started to analyze advanced
metric for player performance analysis that can lead in

- player scouting
- decision making of player’s contract.


## Introduction (2)


While comparing teams or player, the expected goal metric provides a
good statistic in the tactical match analysis. 

The xG model is a probabilistic model that assign score between 0 and 1 from any
observed shot in a match. The model has development using
event-level football data from StatsBomb’s data.

Usually, xG models do not account fot the players
who take the shots, and this assumptions does not seem suitable, 
since the player’s skill could influence the success of shot-
conversion. 

## Aim 


The aim of the analysis is to understand:

- The variables that can influence the shots result
- If the player' skill influence the shot results 

Before going any further on the analysi, it is need to describe the variables that we are going to use for conducting this analysis

##  Dataset

The data are collected by tracking players over the course of
football matchs and recording their actions such shots, passes
or others. For the xG model the data were taken from the
shot event. These data, collected from StatsBomb are spatially manipulated in order
to colecct relevant informations of the shots



## Data extrapolation


![Messi's Goal against Albacete](img/messi_goal_imf){width=50%}



## Dataset description 

The dataset is composed by 13 variable and 7437 observations that reppresent shots of different $42$ player of Barcelona F.C.

- Three continuous predictors: shot-taker’s distance to goal,
angle of shot and time.
- Three binary predictors: tecnique, first touch, inside 18
metres
- Two multi-class predictors: body part and preferred type

## Explanatory Analysis


```{r Bar plot of goal}
barca_shot <- shots[shots$team == "Barcelona",]

barplot(prop.table(table(barca_shot$goal)),
        main = "Bar plot of shots results",
        names.arg = c("No score","Score")
        )

```


## Explanatory Analysis (2)

```{r plot number of goal vs distance}
barca_shot$goal <- as.integer(factor(barca_shot$goal))
distance_class <- factor(cut(barca_shot$shot_distance , breaks = c(0,5,10,15,20,25,30)))
plot(distance_class[ barca_shot$goal == 1], main = "Number of goal per distance class",
     xlab = "Distance class" , ylab = "Number of goal")
```


## Hierarchical assumptions

A reasonable assumption is that player has a unique type of
shooting. Therefore, given a player, the shots are inerehntly
correlate to each each other and behave indipendently form the
shot of another player. This introduces within-player correla-
tions (event-level data is longitudinal data). As a result, every
shot is essentially grouped or nested under a player. Therefore,
models need to be fed information about the hierarchical struc-
ture, otherwise it may lead to biased infe


## Example of the hierarchical structure



![Example of Hierarchical structure](img/messi_inista){width=50%}



## Model specification


For targeting the target variable, two models are choosen.

- GLM (Generalize linear model):  do not fed information about the hierarchical structure
- GLMM (Generalize linar mixed model): do fed information about the hierarchical structure

## Generalized linear model specifications

$$g(\pi_i) = \beta_0 + \beta_1 x_1 + \cdots + \beta_p x_p$$
where:

- $g$ is the logit link function 
- $\pi_i$ is the odds for a goal $\frac{Pr(Y = 1)}{Pr(Y = 0)}$.

## Generalized linear mixed model specifications

In order to fed the information about the players structure a Linear mixed effect model only with a random intercept it is choosen. 

Let us define the equation of the model and some basic notion:


$$g(\pi_{i,j}) = \beta_0 + \beta_1 x_{i,j,1} + \beta_2 x_{i,j,2} \cdots \beta_p x_{i,j,p} + \delta_j $$
\begin{itemize}
    \item $g = 1,2, \dots 42$ number of players
    \item $i = 1 \dots n_g$ Number of shots done by the $g-th$ player
    \item $\pi{_i,g} =  $  odds ratio of score from  the $i-th$ shot and the $g-th$ player
\end{itemize}


## Dimensionality of the parameters

Let us define the dimensionality:

\begin{itemize}
    \item $X_{g,i} \in (k \times 1)$ where $k$ is the number of predictors.
    \item $\underline{\beta} \in (k \times 1)$ can be considered as a vector of fixed effect since it does not depend on the player.
    \item  $\delta{j}$ represent the random effect linked to the g-th player.
    \item The vector $\underline{\epsilon_{g}}$ is the vector of error terms.
\end{itemize}


## Sampling model assumptions

The $\epsilon_g$ $\perp$ $b_g \forall g = 1 \cdots 42$.
The $b_g$ are normally distribuited with their own variance.
The $\epsilon_g$ are normally distribuited.

Conditionally on the group effect, the probability of scoring 
can be considered independent since we are deleting their common factors.

## Poterior Analysis Distribution of the beta coefficient

```{r beta coef , warning=FALSE , echo=FALSE , comment=FALSE, fig.height=5.6}
load("Rdata/model_jags.RData")
# Reshape the dataframe into a long format
# Create the density plot with facet_wrap

#df_long <- reshape2::melt(fixed_beta)
load("Rdata/df_long.RData")


ggplot(df_long, aes(x = value, fill = variable)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ variable, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())


#plot(1:length(fixed_beta$Beta.2),cumsum(fixed_beta$Beta.2 ) / (1:length(fixed_beta$Beta.2)) )
```


## Betas coefficient

\begin{table}[b]
\begin{center}


    
\begin{tabular}{|l|l|}
\hline
Predictors              & Effects \\ \hline
$\beta_{1}$: Intercept                & -1.5087   \\ \hline
$\beta_{2}$: shot\_distance           & -0.184  \\ \hline
$\beta_{3}$: shot\_angle              & 0.0011   \\ \hline
$\beta_{4}$: bodypartLeft Foot        & 1.385   \\ \hline
$\beta_{5}$: bodypartOther Bodypart   & 0.4860   \\ \hline
$\beta_{6}$: bodypartRight Foot       &  1.3539   \\ \hline
$\beta_{7}$: techniqueVolley         & -0.579  \\ \hline
$\beta_{8}$: first\_touchTrue       &  0.1215  \\ \hline
$\beta_{9}$: prefered\_typeLeft Foot & 10.078   \\ \hline
$\beta_{10}$: prefered\_typeRight Foot & 0.523   \\ \hline
$\beta_{11}$: inside\_18True          & -0.615  \\ \hline
$\beta_{12}$: time              & 0.0021   \\ \hline
\end{tabular}
\end{center}

\end{table}



## Player impact on goal

todo
## Model comparison via DIC

\begin{align*}
\begin{cases}
H_0 : \sigma_{b_0} = 0\\
H_1 : \sigma_{b_0} > 0
\end{cases}
\end{align*}


Since the DIC the GLM is lower than the DIC of the GLMM , it is possible to claim that data do not show enough evidence in order to rejcet the null hypothesis.

So we can conlcude that the assumptions of the hierarchical structure does not fit with this data


## Discussion

todo

## References

todo


